---
title: "TidyFriday"
author: "Lisbeth, Kirstine & Marlene"
format: 
 revealjs: 
  code-fold: true
  progress: TRUE
  slide-number: true
  preview-links: auto
  css: styles.css
  embed-resources: true
  smaller: true
  scrollable: true
editor: visual
---

## Bee Colonies Lost {background-color="black" background-image="https://media0.giphy.com/media/v1.Y2lkPTc5MGI3NjExazh3NTdlZ3ptbGJwa2oyeTJ3MjU1aTJoOTkxMTI2MTllYWZiczRkdyZlcD12MV9pbnRlcm5hbF9naWZfYnlfaWQmY3Q9Zw/Yr5cLTZS4Gp3CQRVHs/giphy.webp" background-size="400px"}


## Data and variables {.smaller}  

```{r}
#| echo: true
library(tidyverse)
colony <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-11/colony.csv')

knitr::kable(colony %>% 
  relocate(colony_lost, colony_added, .after = state))
```



## Bullets

When you click the **Render** button a document will be generated that includes:

-   Content authored with markdown
-   Output from executable code

## Code

When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```
